---
title: "A comparison of drm and nlsLM"
author: "Stéphane Laurent"
date: "01/11/2019"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Let's simulate a sample from the 5pl regression model:

```{r data}
A <- 30; D <- 100; B <- 5; xmid <- 55; S <- 25
C <- xmid + log(2^(1/S) - 1)/B 
curve(A + (D-A) / (1 + exp(B*(C-x)))^S, 
      from = 48, to = 62, n = 300, col = "red")
# simulated data
n <- 25
x <- seq(48, 62, length.out = n)
set.seed(3141)
y <- A + (D-A) / (1 + exp(B*(C-x)))^S + rnorm(n, 0, 6)
points(x, y)
```

Now we fit the model with `drc::drm` and `minpack.lm::nlsLM`, assuming the 
parameters $A$, $C$ and $D$ are known.

```{r fit, message=FALSE}
library(drc)
fit_drc <- drm(y ~ x, data = data.frame(x=x, y=y),
           fct = L.5(fixed = c(NA, A, D, C, NA),
                     names = c("minusB", "A", "D", "C", "S")))
#
library(minpack.lm)
fit_minpack <- nlsLM(
  y ~ A + (D-A)/(1 + exp(B*(C-x)))^S,
  data = data.frame(x=x, y=y),
  start = c(B = 1, S = 1),
  lower = c(0, 0),
  control = nls.lm.control(maxiter = 1024, maxfev=10000))
```

Let's look at the estimates:

```{r estimates}
coef(fit_drc)
coef(fit_minpack)
```

They are totally different. 

Let's plot the likelihood:

```{r graph1}
lhd <- function(y, A, B, C, D, S, sigma){
  exp(sum(dnorm(y, A + (D-A)/(1 + exp(B*(C-x)))^S, sigma, log = TRUE)))
}
BandS <- expand.grid(
  B = seq(0.2, 30, length.out = 150),
  S = seq(0.1, 50, length.out = 150)
)
dat <- cbind(BandS, lhd = apply(BandS, 1, function(BS){
  lhd(y, A = A, B = BS[1], C = C, D = D, S = BS[2], sigma = 6)
})) 
library(graph3d)
graph3d(dat, ~B, ~S, ~lhd, type = "surface", keepAspectRatio = FALSE)
```

On the grid we used to do this graph, the maximum of the likelihood is 
attained at:

```{r maxlhd}
dat[which(dat$lhd == max(dat$lhd)), ]
```

So, `nlsLM` well estimates the max-likelihood, while the estimates 
provided by `drm` fall in a flat region of the likelihood.

Let'z zoom:

```{r graph2}
BandS <- expand.grid(
  B = seq(0.2, 6, length.out = 150),
  S = seq(0.1, 50, length.out = 150)
)
dat <- cbind(BandS, lhd = apply(BandS, 1, function(BS){
  lhd(y, A = A, B = BS[1], C = C, D = D, S = BS[2], sigma = 6)
})) 
dat[which(dat$lhd == max(dat$lhd)), ]
graph3d(dat, ~B, ~S, ~lhd, type = "surface", keepAspectRatio = FALSE)
```

This is really a failure of `drm`:

```{r summary_drm}
summary(fit_drc)
```

```{r summary_nlsLM}
summary(fit_minpack)
```

